---
title: "Untitled"
author: "Jeffrey A. Walker"
date: "3/27/2018"
output: html_document
---

```{r mossBar, echo=FALSE, out.width="100%", fig.cap="Bar plot of moss data. Error bars are 1 SEM. Letters indicate statistically significant tests of marginal means pooled over levels of other factor"}
path <- '../data/' # for knit
# path <- 'manuscript/data/' # for console
fn <- paste(path, 'moss2.txt', sep='') # for knit
moss <- fread(fn, stringsAsFactors = TRUE)
# quickie to check against supplement - I get different results (small)
# moss[, .(N=.N, SE=sd(moss_biomass_gain)/sqrt(.N)), by=.(MFW, Precipitation, N_addition)]

moss[, MFW:=factor(MFW, c('C', 'B', 'BS', 'BX'))]
moss[, Precipitation:=factor(Precipitation, c('Frequent', 'Moderate', 'Infrequent'))]

pd <- 0.9
gg_bar <- ggplot(data=moss, aes(x=MFW, y=moss_biomass_gain, fill=Precipitation, group=Precipitation)) +
  stat_summary(fun.data = 'mean_se', geom = c('errorbar'), width = 0.2, position = position_dodge(pd)) +
  stat_summary(fun.y=mean, geom="col", position = position_dodge(pd), color='darkGrey') +
  theme_minimal() +
  ylab('Moss biomass gain (g dry mass per microcosm)') +
  scale_fill_brewer(palette = 'Greys') +
  #coord_cartesian(ylim = 0.50) +
  annotate("text", x=1:4, y=c(.5, .5, .5, .5), label=c('B', 'A', 'AB', 'A')) +
  annotate("text", x=c(1-.25, 1, 1+.25), y=c(.275, .275, .275), label=c('a', 'a', 'b')) + 
  annotate("text", x=c(2-.25, 2, 2+.25), y=c(.4, .4, .4), label=c('a', 'a', 'b')) + 
  annotate("text", x=c(3-.25, 3, 3+.25), y=c(.35, .35, .35), label=c('a', 'a', 'b')) + 
  annotate("text", x=c(4-.25, 4, 4+.25), y=c(.475, .475, .475), label=c('a', 'a', 'b'))

gg_bar
```

```{r make_mossHarrell, echo=FALSE, results='hide'}
# average standard error of raw group values
se_sum <- moss[, .(N=.N, SE=sd(moss_biomass_gain)/sqrt(.N)), by=.(MFW, Precipitation)]
mean_se <- mean(se_sum[, SE])

res <- HarrellPlot(x='MFW', y='moss_biomass_gain', g='Precipitation', data=moss, add_interaction = FALSE, interaction.group = TRUE, contrasts.method = 'revpairwise', contrasts.scaling = 'percent', y_label = ('Moss biomass gain (g dry mass per microcosm)'))
gg_hplot <- res$gg

path <- '../output/' # for knit
#path <- 'manuscript/output/' # for console
image_path <- paste(path, 'moss_Hplot.pdf', sep='') # for knit
pwidth <- 8
pheight <- 6
pdf(image_path, width=pwidth, height=pheight, onefile = TRUE)
grid.arrange(gg_hplot, nrow=1, ncol=1) 
dev.off()

# average SE difference
mean_sed <- mean(res$tables$contrasts[, SE])
```

```{r mossHarrell, echo=FALSE, out.width="90%", fig.cap="Harrell plot of moss linear model results. Error bars are 95\\% confidence intervals"}
path <- '../output/' # for knit
#path <- 'manuscript/output/' # for console
image_path <- paste(path, 'moss_Hplot.pdf', sep='') # for knit
knitr::include_graphics(image_path)

```

In figure \@ref(fig:mossBar), I have redrawn figure 3 from @Kardol_Trophic_2016, which shows treatment means and standard errors for moss biomass in response to different combinations of community complexity and precipitation (the original data are available at https://doi.org/10.5061/dryad.66d5f). The letters, which indicate statistically significant $p$-values for tests of marginal means (pooled across the levels of the other factor), draw the researcher and reader into comparing means using the heights of the bars. The letters are helpful for this because the standard error bars are not especially helpful given the computations necessary to go from the illustrated SEMs to the relevant SEDs.

A major advantage of a Harrell plot (figure \@ref(fig:mossHarrell) over mean-and-error plots with letters or asterisks is that a Harrell plot nudges researchers to focus on modeled effect size and uncertainty instead of classification of results into "significant" and "non-significant" bins. Here, I've scaled the contrasts as percents, since the units of per microcosm are rather arbitrary. Certainly, one can use the 95% confidence limits to mentally bin comparisons into significant and non-significant bins, but the plot itself does not encourage it. Instead, the plot encourages a researcher/reader to consider how the range of values in a CI support, or fail to support, different interpretations of the results. For example, trivial to moderate (50%) reductions in biomass as rain decreases is consistent with the $Moderate - Frequent$ contrast CI, but trivial to moderate increases in biomass as rain decreases are not consistent with the CI. Means with standard errors supplemented with letters or asterisks do not convey this information. A bar plot has other disadvanatages relative to a Harrell plot for these data. For example, the bar plot suggests that the response (moss biomass gain) is a count, or some other variable that can only take positive values but the box/dot plot in the bottom panel of the Harrell plot clearly indicates that moss biomass gain can take negative values.
